---
title: "Los Angeles Lakers 2005-2024 Shot Analysis"
author: "Carter Abitz"
date: "2025-04-02"
output: html_document
self-contained: true
---

```{r include = FALSE}
#Load required packages 
library(devtools)
devtools::install_github("abresler/nbastatR@aba9179ef644f263387c1536d6ddd26104d79cf4")
library(nbastatR)
library(dplyr)
library(ggplot2)
library(gganimate)
library(grid)
library(png)
```

```{r include = FALSE}
#Set up parameters and storage for retrieving shot data
years <- 2005:2024
shots_list <- list()

#Increase buffer size to avoid issues when pulling multiple seasons
Sys.setenv("VROOM_CONNECTION_SIZE" = 131072 * 2)

#Download Lakers shot data season-by-season
for (year in years) {
  shots_year <- teams_shots(teams = "Los Angeles Lakers", seasons = year)
  shots_list[[as.character(year)]] <- shots_year
  Sys.sleep(2)  #Delay prevents request overload
}

#Combine yearly data and preprocess
shots <- bind_rows(shots_list, .id = "season") %>%
  mutate(
    season = as.numeric(season),
    shot_made = ifelse(isShotMade, "Made", "Missed")
  )

#Load and return court background image
draw_court <- function() {
  court_image <- readPNG("at0qmprv.png")
  rasterGrob(court_image, width = unit(1, "npc"), height = unit(1, "npc"), interpolate = TRUE)
}

#Apply custom court theme
court <- function() {
  theme_minimal() +
    theme(
      panel.grid = element_blank(),
      panel.background = element_rect(fill = "white", color = NA),
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      plot.title = element_text(size = 18, face = "bold"),
      legend.position = "right"
    )
}

#Static shot chart for all seasons (used as base for animation)
shot_chart <- ggplot(shots, aes(x = locationX, y = locationY, shape = typeShot)) + 
  annotation_custom(draw_court(), xmin = -321, xmax = 321, ymin = -98, ymax = 898) +
  geom_point(aes(color = shot_made), alpha = 0.5, size = 2) +
  scale_color_manual(values = c("Made" = "purple", "Missed" = "gold")) +
  labs(
    title = "Los Angeles Lakers Shot Chart",
    subtitle = "2005–2024",
    x = "Court X", y = "Court Y", color = "Shot Outcome"
  ) +
  court()

#Animated shot chart by season
animated_shot_chart <- shot_chart +
  transition_states(season, transition_length = 2, state_length = 10) +
  labs(title = "Los Angeles Lakers Shot Chart: {closest_state}", shape = "Shot Type")

#Render animation as GIF
animation <- animate(
  animated_shot_chart,
  nframes = 100,
  fps = 5,
  renderer = gifski_renderer()
)
```

```{r echo =FALSE}
#Display animation
animation
```


•	Shot volume has increased slightly, especially from behind the 3-point line.

•	Shot selection has shifted away from mid-range to more efficient zones: 3s and shots at the rim.

•	Field goal percentage has remained relatively the same, even with increased long-range attempts.

•	Notable surge in 3-point shot attempts beginning in 2013 aligns with NBA trends.


Total shot attempts have recently started picking off. From 2005-2012, the Lakers attempted less shots than they did after. From 2005-2012 there were an average of 6625 shots per season. From 2013-2024 there were an average of 7000 shots per season. From 2005-2012, the Lakers shot 1568 3PT shots per season on average. This made up 23.8% of all shot attempts. From 2013-2024, the Lakers shot 2105 3PT shots per season on average. This made up 30.3% of all shot attempts.

From 2005-2012, the Lakers shot 2137 midrange jump shots per season on average. This made up 32.5% of all shot attempts. From 2013-2024, the Lakers shot 1674 midrange jump shots per season on average. This made up 24.1% of all shot attempts. In 2013–2024, the Lakers attempted high-volume 3s. From the 2011-2012 season to the 2012-2013 season, there was a drastic increase in the percentage of 3PT attempts. In the 2011-2012 season, there were 5318 shot attempts and 1111 3PT attempts, resulting in 20.9% of total shots being 3PT’s. In the 2012-2013 season, there were 6632 shot attempts and 2007 3PT attempts, resulting in 30.3% of total shots being 3PT’s. By 2024, 3PT shots made up over 35% (35.8%) of all shot attempts. 

In 2005-2012, the Lakers had an average of 46% field goal percentage. In 2013-2024, the Lakers had an average of 46.2% field goal percentage. Even with the shot attempts becoming more and more centered around 3PT attempts, the field goal percentage has stayed high. 

The Lakers consistently generated high-percentage looks near the basket, with consistent rim pressure. From 2005-2012, the Lakers shot 2974 close range shots (0-9 ft) per season on average. This made up 45.2% of all shot attempts. From 2013-2024, the Lakers shot 3194 close range shots per season on average. This made up 45.8% of all shot attempts. 